### 2024-05-30 2024-5-13/sections/heuristic.tex
change experiments
optimization for nested counting

In the previous paper \cite{Denghang2023}, we naively unwind the counting operators when they are in other counting or complement operator. However, this approach is not efficient when the counting bounds are large. In this section, we present some heuristics to solve the RECL constraints more efficiently.

\subsection{Heuristic for RECL Constraints with Nested Counting}\label{heuristic:nested}
 The naive approach for nested counting in the paper \cite{Denghang2023} is to unwind all inner counting operators. This approach may generate a large number of states, which is linear to the product of the counting bounds in regex and exponential to the length of the regex. For example, naive unwinding regex $(a^{\{1,1000\}})^{\{1,2\}}$ generate a CEFA containing 1000 states. But if we only unwind the outer counting operator of the regex, not inner, we obtain a CEFA containing only 2 states. The side effect is that we need two registers to store the information of conuting in the inner regex (unwinding of outer counting operator repeat the inner regex, which also repeat the registers of the inner regex).We present a heuristic to reduce the number of states generated by the unwinding process. The idea is to unwind the counting operator with less bound and less side effect firstly. We present the heuristic in the following algorithm.

@@

\subsection{Heuristic for RECL Constraints with Complement on Counting}
As we mentioned in the above section \ref{heuristic:nested}, the unwinding process of the complement operator is exponential to the counting bounds in the worst case. We present a heuristic to reduce the number of states generated by the unwinding process. The idea is to under-approximate and over-approximate the language of the regex first, so that we can avoid unwinding counting operators in complement. We present the heuristic in the following algorithm.

>>>>

%!TEX root = ../main.tex

In Section~\ref{sec:algorithm}, to deal with the non-register-representable RECL constraints, we unfold all the non-register-representable occurrences of counting operators. Nevertheless, 
the naive unfolding incurs an exponential blow-up over the formula sizes, especially when the counting bounds are large. 
Recall that an occurrence of counting operators is said to be non-register-representable if it is in the scope of some Kleene star, another counting operator, complement, or language difference. Since Kleene star can be seen as a special case of the counting operators, and language difference $e_1 \setminus e_2$ can be rewritten as $e_1 \cap \overline{e_2}$, without loss of generality, we can assume that a non-register representable occurrence of a  counting operator is in the scope of another counting operator or some complement operator. 
In the sequel, to alleviate the formula-size blow-up problem resulted from the naive unfolding, we present various optimizations, for the nesting of counting operators and the nesting of a counting operator and a complement operator respectively. 
% two situations that a counting operator is in the scope of another counting operator and in the scope of a complement operator respectively.

\subsection{Optimizations for the nesting of counting operators}\label{heuristic:nested}

In the sequel, we propose an optimization for the nesting of counting operators. 
The optimization is designed for the special case that the nesting depth of counting operators is one. For a RECL constraint whose nesting depth of counting operators is greater than two, we can apply the optimization for the inner-most nesting of counting operators and decrement the nesting depth, then apply the optimization again to the resulting constraint, and continue this process, until obtaining a register-representable RECL constraint.

The main idea of the optimization is illustrated by the constraint $x \in (a^{\{1,1000\}})^{\{1,2\}}$. If we apply the naive unfolding in Section~\ref{sec:algorithm}, we will unfold  $a^{\{1,1000\}}$ and obtain the constraint $x \in (a(\varepsilon + a + aa + \cdots + \underbrace{a\cdots a}_{999}))^{\{1,2\}}$. It is easy to observe that a smarter way is to unfold  $^{\{1,2\}}$, instead of $^{\{1, 1000\}}$.  If we do this, then we get a constraint $x \in (a^{\{1,1000\}})(\varepsilon + a^{\{1,1000\}})$, whose size is much smaller, compared to $x \in (a(\varepsilon + a + aa + \cdots + \underbrace{a\cdots a}_{999}))^{\{1,2\}}$. (Recall that we assume a binary encoding of the integer constants in counting operators.)

Let us describe the optimization more precisely. For a regular expression $x \in (e_1)^{\{m,n\}}$ where $e_1$ contains counting operators (but does not contain the nesting of counting operators), we can unfold the counting operators in the following ways to obtain a register-representable RECL constraint, (i) unfold the counting operators in $e_1$, (ii) unfold the counting operator $^{\{m,n\}}$. Let $e', e''$ denote the resulting constraints. We estimate the sizes of the two CEFAs constructed from $e', e''$ respectively, say $N'$ and $N''$. If $N' \le N''$, then we choose to unfold the counting operators in $e_1$. Otherwise, we choose to unfold the counting operator $^{\{m,n\}}$. Note that when we estimate the sizes of the CEFAs, we do not really need unfold the counting operators. We can simply compute the estimations of the CEFA sizes, without any unfolding.

\zhilin{stopped here}

In the sequel, we show how we estimate the sizes of CEFAs of the register-representable

The naive approach for nested counting in the paper \cite{Denghang2023} is to unwind all inner counting operators. This approach may generate a large number of states, which is linear to the product of the counting bounds in regex and exponential to the length of the regex. For example, naive unwinding regex $(a^{\{1,1000\}})^{\{1,2\}}$ generate a CEFA containing 1000 states. But if we only unwind the outer counting operator of the regex, not inner, we obtain a CEFA containing only 2 states. The side effect is that we need two registers to store the information of conuting in the inner regex (unwinding of outer counting operator repeat the inner regex, which also repeat the registers of the inner regex).We present a heuristic to reduce the number of states generated by the unwinding process. The idea is to unwind the counting operator with less bound and less side effect firstly. We present the heuristic in the following algorithm.

@@

\subsection{Optimizations for the nesting of a counting operator and a complement operator}
As we mentioned in the above section \ref{heuristic:nested}, the unwinding process of the complement operator is exponential to the counting bounds in the worst case. We present a heuristic to reduce the number of states generated by the unwinding process. The idea is to under-approximate and over-approximate the language of the regex first, so that we can avoid unwinding counting operators in complement. We present the heuristic in the following algorithm.

### 2024-05-30 2024-5-13/sections/introduction.tex
optimization for nested counting

\smallskip
\noindent
\emph{Organization.} 
The rest of this paper is structured as follows: Section~\ref{sec:overview} gives an overview of the approach in this paper. Section~\ref{sec:pre} introduces the preliminaries. 
Section~\ref{sec:recl} presents the syntax and semantics of RECL. 
Section~\ref{sec:automaton} defines CEFA. Section~\ref{sec:algorithm} introduces the algorithm to solve RECL constraints. Section~\ref{sec:implementation} shows the experiment results. Section~\ref{sec:conclu} concludes this paper.

>>>>

\smallskip
\noindent
\emph{Structure.} 
The rest of this paper is structured as follows: Section~\ref{sec:overview} gives an overview of the approach in this paper. Section~\ref{sec:pre} introduces the preliminaries. 
Section~\ref{sec:recl} presents the syntax and semantics of RECL. 
Section~\ref{sec:automaton} defines CEFA. Section~\ref{sec:algorithm} introduces the algorithm to solve RECL constraints. 
Section~\ref{sec:heuristic} presents some heuristics for improving the performance of solving RECL constraints where the counting operators are inside the scope of some other counting operators or complement operators. \zhilin{double check later}
Section~\ref{sec:implementation} shows the experiment results. Section~\ref{sec:conclu} concludes this paper.

A preliminary version of this paper has appeared in the proceedings of SETTA 2023~\cite{Denghang2023}. This paper extends the SETTA paper in the following two aspects: (i) To improve the performance of complex RECL constraints where the counting operators are nested with some other counting operators or nested with the complement operators, various heuristics are proposed, (ii) moreover, to evaluate the performance of these heuristics, the benchmarks as well as the experiments are extended.   \zhilin{double check later}
